### 2020-10-31 ChapterBackground/chapterBackground.tex
Fixed: Final pass from proof-read

Consistency resolution can divided into the following two categories:
\begin{itemize}
\item Decision-making. Determining the outcome of decisions in the presence of inconsistency. The decisions can be determined by choosing a perspective, i.e. the server, in which the server maintains the whole state necessary to determine the decision; source client, the perspective of the client that initiates the action (e.g. shoots an enemy player); and target client, the perspective of the client that the action targets (e.g. target of an enemy player shooting) \cite{Savery2014Consistency}.

@@

Although there has been research utilising multiple servers to distribute the task of solving physics-based problems (e.g., \cite{mashayekhi2018automatically}), there is no known literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to this challenge is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs), of which online gaming can be considered a subcategory.

>>>>

Consistency resolution can be divided into the following two categories:
\begin{itemize}
\item Decision-making. Determining the outcome of decisions in the presence of inconsistency. The decisions can be determined by choosing a perspective, i.e. the server, in which the server maintains the whole state necessary to determine the decision; source client, the perspective of the client that initiates the action (e.g. shoots an enemy player); and target client, the perspective of the client that the action targets (e.g. target of an enemy player shooting) \cite{Savery2014Consistency}.

@@

Although there has been prior research utilising multiple servers to distribute the task of solving physics-based problems (e.g., \cite{mashayekhi2018automatically}), there is no known literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to this challenge is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs), of which online gaming can be considered a subcategory.

### 2020-10-31 ChapterConclusion/chapterConclusion.tex
Fixed: Final pass from proof-read

The overhead of boundary objects graph (Fig. \ref{fig_PerfOverheadColumn}) contains a lot of noise and the results are close to the error margin of the fitted line (~0.5ms for the number of objects involved). However, it appears as though there is a positive trend between increasing nodes and increasing overhead, which converges at 8 servers. It is also not clear what the exact relationship is between boundary objects and frame-time as the data for 3 and more servers are indicative of a linear relationship, however, the data for 2 servers suggests an exponential relationship, but based only on the point at 2500 objects and this may be a result of noisy data. More data is required to draw a more concrete conclusion, such as a greater number of iterations or experiments that have greater numbers of objects per server when more servers are used.

The optimal scale of the system was explored and it can be concluded in the case of these performance experiments, going up to 10,000 objects, using 8 servers is the optimal amount. 8 servers results in a performance increase by a factor of 4.22 (see Fig. \ref{fig_FittedCoeff}). Increasing servers beyond 8 results in minimal return and the coefficients suggest performance starts to reduce slightly, which may be a result of both overheads and reduced object efficiency (see Fig. \ref{fig_CostPerObjPerNodeColumn}). However, further work should be carried out to determine how the system performs with more than 10,000 objects and with greater numbers of servers.

>>>>

The overhead of boundary objects graph (Fig. \ref{fig_PerfOverheadColumn}) contains a lot of noise and the results are close to the error margin of the fitted line (~0.5ms for the number of objects involved). However, it appears as though there is a positive trend between increasing nodes and increasing overhead, which converges at 8 servers. It is also not clear what the exact relationship is between boundary objects and frame-time as the data for 3 and more servers are indicative of a linear relationship. However, the data for 2 servers suggests an exponential relationship based only on the single data point at 2500 objects and this may be a result of noisy data. More data is required to draw a more concrete conclusion, such as a greater number of iterations or experiments that have greater numbers of objects per server when more servers are used.

The optimal scale of the system was explored, and it can be concluded in the case of these performance experiments, going up to 10,000 objects, using 8 servers is the optimal amount. 8 servers results in a performance increase by a factor of 4.22 (see Fig. \ref{fig_FittedCoeff}). Increasing servers beyond 8 results in minimal return and the coefficients suggest performance starts to reduce slightly, which may be a result of both overheads and reduced object efficiency (see Fig. \ref{fig_CostPerObjPerNodeColumn}). However, further work should be carried out to determine how the system performs with more than 10,000 objects and with greater numbers of servers.

### 2020-10-31 ChapterProblemDefinition/chapterProblemDefinition.tex
Fixed: Final pass from proof-read

Due to this non-determinism, if the collision response is performed on different servers with the same group of objects, they will not lead to the same results. PhysX has an option for increased determinism, but as servers will have different states, as they will be hosting objects that the other server is not, this will not guarantee the exact same results. The results of a single differing collision response can create a significantly different state of objects, for example if a stack of cubes is dropped onto a flat, level plane, the cubes land is such a way where the stack is on the verge of collapsing. In one iteration, the stack of cubes may remain standing, in a second iteration, the stack of cubes may collapse.

@@

\section{Benefits of Proposed Solution}
Aura Projection has the following benefits over other considered solutions. AP guarantees strong consistency of collisions, meaning diverging results on different servers are prevented. Objects are only simulated once on a one server, instead of being simulated on multiple servers, requiring more computation time per object. No synchronisation is required between servers, as each server is running an entirely autonomous simulation, there is no need for any synchronisation to take place between servers.

\section{Limits of Proposed Solution}
Performance of the system will be measured using the highest cost across all servers, this will be referred to as $costmax$. If the addition of one or more servers (and/or re-partitioning of servers) cannot be done without causing a higher $costmax$, the system cannot be scaled up further.

@@

Let $costmax_n$ be the current $costmax$ of system and $costmax_{n+1}$ be the $costmax$ of a potential new configuration.

@@

Example 1:
Let 1 non-boundary object have the cost value of 1 and suppose the B:N ratio is a constant 2:1.
1 server with 1000 non-boundary objects. The $costmax_n$ is 1000. 1 server is added, the simulation is partitioned so that 500 non-boundary objects exist on each. The $costmax_{n+1}$ is 500. 500<1000, so new configuration results in better performance.

Example 2:
2 servers with 1000 non-boundary objects each. The $costmax_n$ is 1000. 1 server is added, the simulation, the best partitioning that can be achieved results in 1 server having 250 non-boundary and 400 boundary objects. The $costmax_{n+1}$ is (400*B:N)+250 = 1050. 1050>1000, so the new configuration results in worse performance. As this is the best partitioning that could be achieved the system cannot be scaled up further.

>>>>

Due to this non-determinism, if the collision response is performed on different servers with the same group of objects, they will not lead to the same results. PhysX has an option for increased determinism, but as servers will have different states, as they will be hosting objects that the other server is not, this will not guarantee the exact same results. The results of a single differing collision response can create a significantly different state of objects, for example if a stack of cubes is dropped onto a flat, level plane, the cubes land in such a way where the stack is on the verge of collapsing. In one iteration, the stack of cubes may remain standing; in a second iteration, the stack of cubes may collapse.

@@

\section{Benefits of Proposed Solution}
Aura Projection has the following benefits over other considered solutions. AP guarantees strong consistency of collisions, meaning diverging results on different servers are prevented. Objects are only simulated once on one server, instead of being simulated on multiple servers, requiring more computation time per object. No synchronisation is required between servers; as each server is running an entirely autonomous simulation, there is no need for any synchronisation to take place between servers.

\section{Limits of Proposed Solution}
Performance of the system will be measured using the highest cost across all servers, which will be referred to as $costmax$. If the addition of one or more servers (and/or re-partitioning of servers) cannot be done without causing a higher $costmax$, the system cannot be scaled up further.

@@

Let $costmax_n$ be the current $costmax$ of the system and $costmax_{n+1}$ be the $costmax$ of a potential new configuration.

@@

Example 1:
Let 1 non-boundary object have the cost value of 1 and suppose the B:N ratio is a constant 2:1.
The example begins with 1 server with 1000 non-boundary objects and the $costmax_n$ is 1000. 1 server is added, and the simulation is partitioned so that 500 non-boundary objects exist on each. The $costmax_{n+1}$ is 500. 500<1000, so the new configuration results in better overall performance of the system.

Example 2:
The example begins with 2 servers with 1000 non-boundary objects each and the $costmax_n$ is 1000. 1 server is added, and the simulation is partitioned. The best partitioning that can be achieved results in 1 server having 250 non-boundary and 400 boundary objects. The $costmax_{n+1}$ is (400*B:N)+250 = 1050. 1050>1000, so the new configuration results in worse performance. As this is the best partitioning that could be achieved the system cannot be scaled up further.

### 2020-10-31 ChapterResults/chapterResults.tex
Fixed: Final pass from proof-read

Experiments were also carried out using servers in a corner layout. The layouts of 3 and 4 servers are shown in Fig. \ref{ServerCorner}, in which the 4 server case is a 2x2 grid with a single corner intersection in the centre. The 9 server case is a 3x3 grid with 4 corner intersections. Injection rates and volume dimensions remain the same as in the column based experiments. Fig. \ref{fig_PerCor} shows the graph illustrating performance.

As discussed in \ref{Topology}, topologies with corner boundaries between server regions have higher communication requirements, this is because servers must communicate with all neighbouring servers. The highest number of expected neighbours for a given server in a grid layout is 8 servers, this applies to the server in the centre of the grid illustrated in Fig. \ref{BrickLayout} and any server not on the edge of the grid. The graph in Fig. \ref{fig_PerCor} demonstrates that increasing the number of servers lowers the frame-time. This remains true even in the presence of corner boundaries and with the maximum number of expected neighbours, 8, in a grid layout (meaning maximum communication requirement). The additional processing overhead of the higher communication requirements (as messages must be exchanged with all neighbouring servers) on the server with 8 neighbours is outweighed by the performance gains of partitioning the simulation using additional servers. In other words, 9 servers in grid layout outperform (have a lower frame-time than) a single server, despite the server in the centre of the grid needing to communicate with the maximum number of neighbouring servers for a grid layout.

The performance cost of the higher communication requirements for corner layouts is reflected in the performance graphs as the respective number of servers in column layout outperform those in corner layout. i.e. 3 servers in column layout have a lower frame-time than 3 servers in corner layout.

@@

\subsection{Object Migrations}
In this sub-section we show the number of migrations of objects between servers. This is done using the average cumulative migrations between servers over time for the different numbers of servers. For each iteration of the experiment, the maximum number of cumulative migrations across all nodes for each time interval is taken, the mean of this value for each time interval is then taken across the 50 repetitions. This was done for both the column (see Fig. \ref{fig_MigrationsColumn}) and corner experiments (see Fig. \ref{fig_MigrationsCorner}).

In the experiments objects are injected at a fixed rate across all servers into random locations within regions near and far from the boundary. As the number of servers increase, the number of objects injected per server decreases. As a result of this, experiments involving more servers will have a lower number and density of objects leading to fewer object migrations.

@@

\subsection{Object Updates}
In the sub-section we show the number of object updates sent between servers. An update is sent from the host server to any neighbours with interest in that object (i.e. when an object is within the aura distance of a boundary), whenever the object's position changes. In order to show the number of updates, for each time interval of 5 seconds, the maximum number of updates is taken across all servers for each iteration of the experiment. The mean of these maximums is then taken for each number of servers and plotted over elapsed time of the experiment. This was done for both the column (see Fig. \ref{fig_UpdatesColumn}) and corner experiments (see Fig. \ref{fig_UpdatesCorner}).

The graphs demonstrate that as elapsed time increases, so does the number of object updates being sent in both the column and corner experiments. However, the relationship is not a linear one. The density of objects around the boundary will affect the number of updates being sent in two ways: (1) Updates are not sent for objects at rest; a higher density of objects means a newly injected object is more likely to collide with an object at rest, resulting in updates needing to be sent for both updates (2) A higher density of objects means a newly injected object is more likely to collide with another object and come to rest sooner than if no object were there.

Comparing the column and corner experiments, the corner experiments have a lower object density in the boundary region as objects are spawned in multiple boundary regions on all servers. As a result of this, the corner experiments result in fewer object updates for all nodes (3, 4 \& 9) compared to the same number of nodes in column layout.

@@

\subsection{Performance Overhead}
In this sub-section we address the performance overhead of Aura Projection. This overhead is calculated by first measuring the performance cost (frame-time) in terms of objects per server (the number of objects is known as the injection rate of objects is known). This is shown in Fig. \ref{fig_PerfPerNodeRaw} and a line can then be fitted to the performance cost of the single server (centralised experiment) experiment with increasing numbers of objects. This provides a baseline cost for non-boundary objects.

@@

The relationship between objects per node and frame-time is approximately exponential and can be seen more clearly when plotting the measurements of frame-time on a $log_{10}$ scale (shown in Fig. \ref{fig_PerfPerNodeRawLog}). The line is fitted using linear regression on the $log_{10}$ of the frame-time (shown in Fig. \ref{fig_PerfFitted}).

@@

The cost of non-boundary objects can then be subtracted from the performance cost of each number of servers using the baseline from the fitted line for the centralised experiment (the ratio of boundary to non-boundary objects is 1:1, so the number of boundary and non-boundary objects is known) leaving just the cost of boundary objects. To calculate the overhead of boundary objects, the cost of the same number of non-boundary objects is subtracted, leaving just the overhead of boundary objects relative to the cost of non-boundary objects (shown in Fig. \ref{fig_PerfOverheadColumn}).

@@

This means, for example, if 1000 objects took 6ms to perform a single physics time-step, 500 objects would take more than half of 6ms. This has important consequences when distributing the workload as it means splitting objects between two servers may not double the performance even without accounting for overhead. Likewise, splitting the objects between two servers could also potentially lead to more than double the performance. For example, if 8000 objects exist on one server and takes 80ms to perform a physics time-step. If the objects are then split across two servers, resulting in 4000 objects on each, it would take approximately 12ms to perform a physics time-step, a 6.67 factor increase in performance (not accounting for overhead).

The relationship between number of objects and frame-time is approximately exponential (shown in Fig. \ref{fig_PerfColumn}). In order to determine the optimal scale for this experiment, we can calculate the coefficients for approximate exponential functions for the performance of each number of servers. Ideally, a double in the number of servers, should lead to a double in performance and therefore double the exponential coefficient. First, approximate functions were fitted for each node using linear regression on the $log_{10}$ of the frame-time (shown in Fig. \ref{fig_FittedCostColumns}). This gives a coefficient for each node. This coefficient can then be used to work out relative performance. 2 servers led to a increase in performance by a factor of 1.74, 4 servers a factor of 3.40 and 8 servers a factor of 4.22. It should be noted that the accuracy of the coefficients for higher numbers of servers (i,e, 8, \& 10) will be low as the relationship appears linear when there are low numbers of objects per server. As a result of this, it is not possible to accurately predict the performance of the system for more than 8 servers with more than 10,000 objects (the total used in this experiment).

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\textwidth]{FittedCostColumns}
\caption{Fitted Cost of Increasing Objects (in column layout). The shapes indicate the results from the experiment, the lines indicate the fitted functions}
\label{fig_FittedCostColumns}
\end{figure}

The coefficients used in the approximate functions can then be plotted and compared to the ideal line, show in Fig. \ref{fig_FittedCoeff}. A line was then fitted using a self-starting asymptotic regression model to the curve of coefficients, to approximate the point at which the performance factor increases converge. The result of which is 4.26.

\begin{figure}[!t]
	\centering
	\includegraphics[width=\textwidth]{FittedCoeff}
	\caption{Fitted Cost of Increasing Objects (in column layout). Coefficients used in Fig. \ref{fig_PerfColumn} are in red, the ideal line of Y=X (i.e. doubling servers, doubles performance) is in green and a line fitted on the values of the coefficients is in blue.}
	\label{fig_FittedCoeff}
\end{figure}
\clearpage

@@

The first experiment conducted was the varying speed experiment. In this experiment the speed of each of the two objects was varied from $1m\mathord{\cdot}s^{-1}$ to double the the following tolerances: $16m\mathord{\cdot}s^{-1}$; $32m\mathord{\cdot}s^{-1}$ and $48m\mathord{\cdot}s^{-1}$. Increments of $1m\mathord{\cdot}s^{-1}$ were used and the whole experiment repeated $50$ times. In these experiments it is assumed 1 unit is 1 metre. The speed tolerances chosen represent categories of typical fast moving objects in games. The categories selected were: slow-moving vehicles ($16m\mathord{\cdot}s^{-1}$); fast-moving/motorway-speed vehicles ($32m\mathord{\cdot}s^{-1}$); and racing vehicles ($48m\mathord{\cdot}s^{-1}$).

>>>>

Experiments were also carried out using servers in a corner layout. The layouts of 3 and 4 servers are shown in Fig. \ref{ServerCorner}, in which the 4 server case is a 2x2 grid with a single corner intersection in the centre. The 9 server case is a 3x3 grid with 4 corner intersections. Injection rates and volume dimensions remain the same as in the column-based experiments. Fig. \ref{fig_PerCor} shows the graph illustrating performance.

As discussed in \ref{Topology}, topologies with corner boundaries between server regions have higher communication requirements. This is because servers must communicate with all neighbouring servers. The highest number of expected neighbours for a given server in a grid layout is 8 servers, which applies to the server in the centre of the grid illustrated in Fig. \ref{BrickLayout} and any server not on the edge of the grid. The graph in Fig. \ref{fig_PerCor} demonstrates that increasing the number of servers lowers the frame-time. This remains true even in the presence of corner boundaries and with the maximum number of expected neighbours, 8, in a grid layout (meaning maximum communication requirement). The additional processing overhead of the higher communication requirements (as messages must be exchanged with all neighbouring servers) on the server with 8 neighbours is outweighed by the performance gains of partitioning the simulation using additional servers. In other words, 9 servers in grid layout outperform (have a lower frame-time than) a single server, despite the server in the centre of the grid needing to communicate with the maximum number of neighbouring servers for a grid layout.

The performance cost of the higher communication requirements for corner layouts is reflected in the performance graphs as the respective number of servers in column layout outperform those in corner layout, i.e. 3 servers in column layout have a lower frame-time than 3 servers in corner layout.

@@

\subsection{Object Migrations}
In this sub-section we show the number of migrations of objects between servers. This is done using the average cumulative migrations between servers over time for the different numbers of servers. For each iteration of the experiment, the maximum number of cumulative migrations across all nodes for each time interval is taken and the mean of this value for each time interval is then taken across the 50 repetitions. This was done for both the column (see Fig. \ref{fig_MigrationsColumn}) and corner experiments (see Fig. \ref{fig_MigrationsCorner}).

In the experiments, objects are injected at a fixed rate across all servers into random locations within regions near and far from the boundary. As the number of servers increases, the number of objects injected per server decreases. As a result of this, experiments involving more servers will have a lower number and density of objects leading to fewer object migrations.

@@

\subsection{Object Updates}
In this sub-section we show the number of object updates sent between servers. An update is sent from the host server to any neighbours with interest in that object (i.e. when an object is within the aura distance of a boundary), whenever the object's position changes. The number of updates shared between servers will be shown in graphs produced using the following process: for each time interval of 5 seconds, the maximum number of updates is taken across all servers for each iteration of the experiment. The mean of these maximums is then taken for each number of servers and plotted over elapsed time of the experiment. This was done for both the column (see Fig. \ref{fig_UpdatesColumn}) and corner experiments (see Fig. \ref{fig_UpdatesCorner}).

The graphs demonstrate that as elapsed time increases, so does the number of object updates being sent in both the column and corner experiments. However, the relationship is not a linear one. The density of objects around the boundary will affect the number of updates being sent in two ways: (1) Updates are not sent for objects at rest; a higher density of objects means a newly injected object is more likely to collide with an object at rest, resulting in updates needing to be sent for both objects (2) A higher density of objects means a newly injected object is more likely to collide with another object and come to rest sooner than if no object were there.

Comparing the column and corner experiments, the corner experiments have a lower object density in the boundary region as objects are spawned in multiple boundary regions on all servers. As a result of this, the corner experiments result in fewer object updates for all numbers of servers (3, 4 and 9) compared to the same number of nodes in column layout.

@@

\subsection{Performance Overhead}
In this sub-section we address the performance overhead of Aura Projection. This overhead is calculated by first measuring the performance cost (frame-time) in terms of objects per server (the number of objects is known because the injection rate of objects is known). This is shown in Fig. \ref{fig_PerfPerNodeRaw}. A curve can then be fitted to the performance cost of the single server (centralised) experiment with increasing numbers of objects. This provides a baseline cost for non-boundary objects.

@@

The relationship between objects per node and frame-time is approximately exponential and can be seen more clearly when plotting the measurements of frame-time on a $log_{10}$ scale (shown in Fig. \ref{fig_PerfPerNodeRawLog}). The curve is fitted using linear regression on the $log_{10}$ of the frame-time (shown in Fig. \ref{fig_PerfFitted}).

@@

The cost of non-boundary objects can then be subtracted from the performance cost of each number of servers using the baseline from the fitted curve for the centralised experiment (the ratio of boundary to non-boundary objects is 1:1, so the number of boundary and non-boundary objects is known) leaving just the cost of boundary objects. To calculate the overhead of boundary objects, the cost of the same number of non-boundary objects is subtracted, leaving just the overhead of boundary objects relative to the cost of non-boundary objects (shown in Fig. \ref{fig_PerfOverheadColumn}).

@@

This means, for example, if 1000 objects took 6ms to perform a single physics time-step, 500 objects would take more than half of 6ms. This has important consequences when distributing the workload as it means splitting objects between two servers may not double the performance even without accounting for overhead. Likewise, splitting the objects between two servers could also potentially lead to more than double the performance. For example, if 8000 objects exist on one server and the simulation takes 80ms to perform a physics time-step, then if the objects are split across two servers, resulting in 4000 objects on each, it would take approximately 12ms to perform a physics time-step. This would be a 6.67 factor increase in performance (not accounting for overhead).

The relationship between number of objects and frame-time is approximately exponential (shown in Fig. \ref{fig_PerfColumn}). In order to determine the optimal scale for this experiment, we can calculate the coefficients for approximate exponential functions for the performance of each number of servers. Ideally, a double in the number of servers should lead to a double in performance and therefore double the exponential coefficient. First, approximate functions were fitted for each node using linear regression on the $log_{10}$ of the frame-time (shown in Fig. \ref{fig_FittedCostColumns}). This gives a coefficient for each node. This coefficient can then be used to work out relative performance. 2 servers led to an increase in performance by a factor of 1.74, 4 servers a factor of 3.40 and 8 servers a factor of 4.22. It should be noted that the accuracy of the coefficients for higher numbers of servers (i.e. 8, \& 10) will be low as the relationship appears linear when there are low numbers of objects per server. As a result of this, it is not possible to accurately predict the performance of the system for more than 8 servers with more than 10,000 objects (the total used in this experiment).

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\textwidth]{FittedCostColumns}
\caption{Fitted Cost of Increasing Objects (in column layout). The shapes indicate the results from the experiment, the curves indicate the fitted functions}
\label{fig_FittedCostColumns}
\end{figure}

The coefficients used in the approximate functions can then be plotted and compared to the ideal line, show in Fig. \ref{fig_FittedCoeff}. A curve was then fitted using a self-starting asymptotic regression model to the curve of coefficients, to approximate the point at which the performance factor increases converge. The result of this is 4.26.

\begin{figure}[!t]
	\centering
	\includegraphics[width=\textwidth]{FittedCoeff}
	\caption{Fitted Cost of Increasing Objects (in column layout). Coefficients used in Fig. \ref{fig_PerfColumn} are in red, the ideal line of Y=X (i.e. doubling servers doubles performance) is in green and a curve fitted on the values of the coefficients is in blue.}
	\label{fig_FittedCoeff}
\end{figure}
\clearpage

@@

The first experiment conducted was the varying speed experiment. In this experiment the speed of each of the two objects was varied from $1m\mathord{\cdot}s^{-1}$ to double the the following tolerances: $16m\mathord{\cdot}s^{-1}$; $32m\mathord{\cdot}s^{-1}$ and $48m\mathord{\cdot}s^{-1}$. Increments of $1m\mathord{\cdot}s^{-1}$ were used and the whole experiment repeated $50$ times. In these experiments, it is assumed 1 unit is 1 metre. The speed tolerances chosen represent categories of typical fast-moving objects in games. The categories selected were: slow-moving vehicles ($16m\mathord{\cdot}s^{-1}$); fast-moving/motorway-speed vehicles ($32m\mathord{\cdot}s^{-1}$); and racing vehicles ($48m\mathord{\cdot}s^{-1}$).
